\thetitle

### 2021-03-02 writing/challenges.tex
touch up in challenges (section 4) and opening of rational (section 5)

However, three major challenges remain:  the isolation of the effect of
blame on the rational programmer across three different mechanisms; the
collection of a large number of representative debugging scenarios;  and a
proper sampling strategy for the huge space of possibilities. A
coincidental challenge is the disparity of gradually typed languages. To
eliminate this variable from the comparison, this paper uses Racket, which
is thus far the only language in which all three major soundness
mechanisms are available in a reasonably robust and directly comparable
manner; for the implementation of the transient semantics, see the
parallel submission of an experience report~\cite{ttt21}.

The {\em first challenge\/} stems from the interplay between blame and the
run-time checks of a gradual type systems. When a check fails the Natural
and Transient semantics assign blame and discard  the information in the
exception from the run-time check. Nevertheless, the exception
information may be equally or even more useful to the rational programmer
than blame. If this is the case blame may not play a critical role for the
rational programmer. Indeed, precisely because they do not
account for confounding factors, ~\citet{lksfd-popl-2020} cannot draw any
conclusions about the pragmatics of blame despite advertisements for the
opposite. Their experiment can only conclude that so-called blame-shifting
works but not that this is because of blame.

@@

Modes also offer a common vocabulary for the differences among the blame
assignment of the three semantic variants.  While Natural assigns blame to
{\em one\/} component, Transient assigns blame to a sequence of
components. The Erasure semantics does not blame a component {\it per
se\/}, but it comes with an exception location and a stack trace, which
implicitly suggests parties to blame.  The different strategies greatly
affect the behavior of the rational programmer (and real programmers,
too).

@@

The {\em second challenge\/} is to find a representative collection of
programs with impedance mismatches. The latter must represent mistakes that programmers
accidentally create and that the run-time checks of academic systems catch. In
other words, the experiment calls for a collection of type mistakes in
mix-typed programs that is representative of those ``in the wild.''
Unfortunately no such collection exists, and with good reason. The kind of
mistakes needed are typically detected by unit or integration tests; even if it
takes some time to find their sources, these mistakes do not make it into code
repositories with appropriate commit messages.

@@

The {\em third challenge\/} is the explosive number of debugging scenarios that
result from the combination of mutation-based scenario generation and mode-based
analysis. All three factors---three different gradual typing systems, the large
number of mutants, and the number of debugging modes---contribute possibly
useful experimental data in a multiplicative manner. Hence, carried out naively,
the experiment would demand an unacceptably large amount of computational
resources.  A practical execution has no option but to {\em sample the space of
scenarios\/}, carefully ensuring that it still produces reproducible
conclusions.

>>>>

However, three major challenges remain:  the isolation of the effect of
blame on the rational programmer across three different mechanisms; the
collection of a large number of representative debugging scenarios;  and a
proper sampling strategy for the huge space of possibilities. A
coincidental challenge is the disparity of the implementations of
gradually typed languages. To eliminate this variable from the comparison,
this paper uses Racket, which is thus far the only language in which all
three major soundness mechanisms are available in a reasonably robust and
directly comparable manner; for the implementation of the transient
semantics, see the parallel submission of an experience
report~\cite{ttt21}.

The {\em first major challenge\/} stems from the interplay between blame and the
run-time checks of a gradual type systems. When a check fails the Natural
and Transient semantics assign blame and discard  the information in the
exception from the run-time check. Nevertheless, the exception
information may be equally or even more useful to the rational programmer
than blame. If this is the case blame may not play a critical role for the
rational programmer. Indeed, precisely because they do not
account for confounding factors, ~\citet{lksfd-popl-2020} cannot draw any
conclusions about the pragmatics of blame despite advertisements for the
opposite. Their experiment can only conclude that so-called blame-shifting
works but not that this is because of blame.

@@

Modes also offer a solution to a derivative challenge; reconciling the
differences among the blame assignment mechanisms of the three semantic variants.
While Natural assigns blame to {\em one\/} component, Transient assigns
blame to a sequence of components. The Erasure semantics does not blame a
component {\it per se\/}, but it comes with an exception location and a
stack trace, which implicitly suggests parties to blame.  The different
strategies greatly affect the behavior of the rational programmer (and
real programmers, too).

@@

The {\em second major challenge\/} is to find a representative collection of
programs with impedance mismatches. The latter must represent mistakes that programmers
accidentally create and that the run-time checks of academic systems catch. In
other words, the experiment calls for a collection of type mistakes in
mix-typed programs that is representative of those ``in the wild.''
Unfortunately no such collection exists, and with good reason. The kind of
mistakes needed are typically detected by unit or integration tests; even if it
takes some time to find their sources, these mistakes do not make it into code
repositories with appropriate commit messages.

@@

The {\em third major challenge\/} is the explosive number of debugging scenarios that
result from the combination of mutation-based scenario generation and mode-based
analysis. All three factors---three different gradual typing systems, the large
number of mutants, and the number of debugging modes---contribute possibly
useful experimental data in a multiplicative manner. Hence, carried out naively,
the experiment would demand an unacceptably large amount of computational
resources.  A practical execution has no option but to {\em sample the space of
scenarios\/}, carefully ensuring that it still produces reproducible
conclusions.

### 2021-03-02 writing/rational.tex
touch up in challenges (section 4) and opening of rational (section 5)

Hence the research problem is how to make the actions of a rational programmer
comparable across all semantics. Sections~\ref{sub:natural}
through~\ref{sub:erasure} describe the various mode of the rational programmer and explain
to what extent they are comparable. Section\ref{sub:experiment}
enhances comparability with the programmer-effort dimension. Finally, the last
subsection states the experimental questions and gives an overview of the
experimental process.

>>>>

Hence the research problem is how to make the modes of the rational programmer
comparable even when they correspond to different semantics. Sections~\ref{sub:natural}
through~\ref{sub:erasure} describe the various modes of the rational programmer and explain
to what extent they are comparable. Section\ref{sub:experiment}
enhances comparability with the programmer-effort dimension. Finally, the last
subsection states the experimental questions and gives an overview of the
experimental process.
